$p$-adic precision

### 2014-02-20 padicprec.tex
Work on Section 3

We first note that a LU decomposition of a given matrix $M$ is unique
if you require further that the $L$-part of this decomposition is
unipotent (\emph{i.e.} all diagonal entries equal $1$). Let us write
$L(M) U(M)$ the LU factorization of $M$ satisfying the above extra
condition (when it exists). There exists Cramer-type formulas that
gives each individual entries of $L(M)$and $U(M)$ in a closed form
(see \cite{caruso:12a	}).

If $M$ is random square matrix over $\Zp$ of side $d$ whose every
coefficient is known up to precision $O(p^N)$, the two following
results are proved in \cite{caruso:12a}:
\begin{itemize}
\item using usual Gauss elimination and tracking $p$-adic precision 
step-by-step, the smallest precision on an entry of $L(M)$ is about
$O(p^{N - \frac{2d}{p-1}})$ on average;
\item computing $L(M)$ by evaluating Cramer-type formulas yields a
result whose every entry is known up to precision $O(p^{N - 2 \log_p 
d})$.
\end{itemize}
If $d$ is large compared to $p$, the second precision is much more
accurate than the first one. This shows once again that ``artifical''
losses of precision appear when one performs Gauss elimination by
tracking precision step-by-step.

@@

\begin{deftn} \label{deftn : diff}
Let $V, W$ be two normed $K$-vector spaces, let $U \subset V$ be an open 
subset of $V$ and let $f : U \rightarrow W$ be a map. Then $f$ is called 
\emph{differentiable} at $v_0 \in U$ if there exists a 
continuous linear map $f'(v_0) : U \rightarrow W$ such that for any 
$\varepsilon >0$, there exists an open neighborhood $U_\varepsilon 
\subset U$ containing $v_0$ with
\[ 
\Vert f(v)-f(w)-f'(v_0) \cdot \left( v-w \right) \Vert 
\leq \varepsilon \Vert v-w \Vert. 
\]
for all $v, w \in U_\varepsilon$.  The linear map $f'(v_0)$ is called the \emph{differential} of $f$ at $v_0$.
\end{deftn}

@@

\begin{proof} % Call it the Banach-Schauder theorem?
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, the open mapping theorem provides a $C>0$ such that 
\[
B_F(1) \subset C  f'(0)(B_E(1)).
\]
Let $\varepsilon>0$ be such that 
$\varepsilon C < \rho$, and choose $U_\varepsilon \subset E$ as in Definition 
\ref{deftn : diff}.  We may assume $U_\varepsilon = B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$ and suppose $H$ is a lattice with $B^-_E(\rho r) \subset H \subset B_E(r).$ 
We seek to show that $f$ maps $H$ surjectively onto $f'(0) (H)$. We first prove 
that $f(H) \subset f'(0) (H)$. Suppose $x \in H$; we first show that $f(x) \in f'(0)(H)$.  By differentiability at 
$0$, $\Vert f(x)-f'(0)(x) \Vert \leq \varepsilon \Vert x \Vert $. 
Setting $y=f(x)-f'(0)(x)$, we have $\Vert y \Vert \leq 
\varepsilon r$. The definition of $C$ implies that $B_F(\varepsilon r) 
\subset \varepsilon C f'(0) (B_E(r))$. Thus there exists $x' \in 
\varepsilon C \cdot B_E(r)$ such that $f'(0) (x') =y$. Since 
$\varepsilon C < \rho$, we get $x' \in B^-_E(\rho r) \subset H$ and 
then $f(x)= f'(0) (x-x') \in f'(0) (H)$.

@@

\todofor{Tristan}{Improve this Subsection}

Before discussing applications of Lemma \ref{lem:main} to ultrametric 
precision, we discuss the case that $f$ is locally analytic.  In this context
we prove that we can recover the constant $\delta$ appearing in the lemma
from the Newton polygon determined by the Taylor series expansion of $f$.

Schneider \cite{schneider:11a}*{\S 6} defines locally analytic functions on finite dimensional vector spaces.
His definition extends easily to $K$-Banach spaces:

\begin{deftn} \label{deftn:analyticNP}
Let $E$ and $F$ be $K$-Banach spaces, and let $U \subset E$ be an open set.
A function $f \col U \rightarrow F$ is said to be \emph{locally analytic} if one may find
an open set $U_x \subset E$ for each $x \in U$ satisfying the following condition.
For each $n \ge 1$ there exists an $n$-linear mapping
$L_n \col E^n \rightarrow F$ so that 
\begin{enumerate}
\item the sum $\sum_{n \geq 1} L_n(y,\ldots,y)$ converges in F for every $y \in U_x,$ 
\item $f(x+y)=f(x)+\sum_{n \geq 1} L_n(y,\dots,y).$
\end{enumerate}
\end{deftn}

We then mostly study such a serie defined at $x$.

\begin{deftn} \label{deftn :  serie and norm}
Suppose that $E$ and $F$ are $\overline{K}$-Banach spaces and $f_n \col E \to F$ are maps satisfying
$\lvert f_n(\lambda x) \rvert = \lvert \lambda \rvert \cdot \lvert f_n (x) \rvert$ for all $\lambda \in \overline{K}$ and $x \in E$.
We define $\lVert f_n \rVert = \sup_{x \in B_E(1)} \vert f_n(x) \vert$ and further assume that all the $\lVert f_n \rVert$ are finite.
Write $f = \sum_n f_n$ and define $\Lambda(f)(v) \in \R \cup \{\infty\}$ for each $v \in \R$ by
\[
\Lambda(f)(v) = \log \sup_{x \in B_E(e^v)} \vert f(x) \vert.
\]
Finally, we define the Newton polygon $\NP(f)$ of $f$ to be the lower convex hull in $\mathbb{R}^2$
of the points $(k, - \log \lVert f_k \rVert )$ for $k \in \Z_{\ge 0}.$ 
\end{deftn}

We define $v_\infty \in \mathbb{R} \cup \left\lbrace \infty \right\rbrace$ to be the limit of the slopes of $\NP(f)$, and
denote by $\NP(f)^*$ the convex conjugate of $\NP(f)$.  The following proposition is a classical consequence of convex analysis.

\begin{prop} \label{prop : conjugate Newton}
Considered as a function on $\R$, $\NP(f)^*$ is defined over $\left] -\infty,v_{\infty} \right[$, though it may be defined for larger values.
\end{prop}

We can interpret the function $\Lambda(f)$ in terms of the conjugate Newton polygon:

\begin{prop} \label{prop : connexion sup Newton}
For any $v \in \left]-\infty, v_\infty \right[,$ \[ \Lambda (f) (v) = \NP(f)^*(v).\]
\end{prop}
\begin{proof}
We prove the equality first for $v$ that are not slopes of $\NP(f)$.  Assume that $v < v_\infty$ is such a slope and set $u = \NP(f)^*(v)$.
Then the intersection of $\NP(f)$ with the line defined by $y=vx-u$ is a vertex of $\NP(f),$ say
$(n, - \log \lVert f_n \rVert )$ for some $n \geq 0$ with $u = vn + \log \lVert f_n \rVert.$

Since $v$ is not a slope of $\NP(f)$, the difference quotients of $\NP(f)$ at $n$ are increasing.  Therefore, we may find a $c>0$ so that $vm-u \leq - \log \lVert f_m \rVert - c \vert n-m \vert$ for any $m\geq 0$.  Since $vm-u=vm-vn-\log \lVert f_n \rVert,$ we get $-vn-\log \lVert f_n \rVert + c \vert n-m \vert  \leq -vm-\log \lVert f_m \rVert .$

For any $x \in  B_E(e^v)$ and $m \geq 0,$ we have
\[
\vert f_m (x) \vert \leq \lVert f_m \rVert e^{vm}
\]
by the definition of $\lVert f_m \rVert$. Therefore
\[
\vert f_m (x) \vert \leq e^{-c \vert n-m \vert } \lVert f_n \rVert  e^{vn}< \lVert f_n \rVert e^{vn}.
\]
Thus, the series converges and by ultrametricity $ \vert f(x) \vert \leq \vert f_n \vert e^{vn},$ and $\Lambda (f)(v) \leq \log \left( \vert f_n \vert e^{vn} \right).$

Now, the definition of $\lVert f_n \rVert$ provides a sequence $(x_i)_{i\geq 0}$ in $B_E(e^v)$ such that $\lim \vert f_n (x_i) \vert= \lVert f_n \rVert e^{vn}.$

Since for any $i,$ $\vert f_m (x_i) \vert \leq e^{-c \vert n-m \vert } \vert f_n \vert  e^{vn},$ then for $i$ large enough, $\vert f_m (x_i) \vert < \vert f_n (x_i) \vert $ and $\vert f(x_i) \vert = \vert f_n (x_i) \vert.$

As a consequence, $\lim_i \vert f(x_i) \vert = \lVert f_n \rVert e^{vn},$ and  $\Lambda (f)(v) = \log ( \lVert f_n \rVert e^{vn}) = u= \NP(f)^*(v).$

Therefore, for any $v \in \left] -\infty,v_{\infty} \right[,$ with $v$ not a slope of $\NP(f),$ we get $\Lambda (f)(v) =\NP(f)^*(v).$
Since $\Lambda (f)$ is increasing, $\NP(f)^*$ is continuous (since it is convex), and $\NP(f)$ has only finitely many slopes smaller than $v_\infty-\epsilon$, for any $\epsilon >0$, we deduce that both functions coincide on $v \in \left] -\infty,v_{\infty} \right[,$
\end{proof}

\begin{lem} Let $f(x)= \sum_{n \geq 0} f_n (x)$ and $g(x)= \sum_{n \geq 0} g_n (x)$ be two series as in definition \ref{deftn :  serie and norm}. Then
\begin{align*}
\Lambda(f+g) &\leq \max (\Lambda(f),\Lambda(g)) \\
\Lambda(f \times g) &\leq \Lambda(f)+\Lambda(g) \\
\Lambda(f \circ g) &\leq \Lambda(f) \circ \Lambda(g)
\end{align*}
\end{lem}

The following classical result from convex analysis will play a crucial role in bounding $\NP(f),$  $\NP(f)^*$ and $\Lambda(f)$.

\begin{prop} \label{prop : legendre transform}
Slopes of $f^*$ corresponds to extreme points of $f$.
\end{prop}

\begin{cor} \label{cor : trunc series}
Let $f(x)= \sum_{n \geq 0} f_n (x)$ be a serie as in definition \ref{deftn :  serie and norm}, $n_0 \geq 0$ and $g(x)= \sum_{n \geq n_0} f_n (x).$ Then $\Lambda(g) \leq \Lambda(f)$ and the slopes of $\Lambda(g)$ are bigger than $n_0$. 
As a consequence, a minoration of $\Lambda(g)$ can be obtain by extending $\Lambda(f)(v)$ with a slope $n_0$ for $v \leq n_0$.
\end{cor}

Now, let $E$ and $F$ be two complete normed $K$-vector spaces, and let $f$ and the $L_k$'s be as in definition \ref{deftn:analyticNP}. We assume $L_1$ is surjective. We define, the $\overline{L_k},$ $\overline{E}$ and $\overline{F}$ by extending scalars to $\overline{K}.$ We define for any $n \geq 1$, $f_n(x)=\overline{L_n}(x,\dots,x)$ for any $x \in \overline{E}.$ 
Let $C$ be given by the open mapping theorem for $L_1$, $\rho \in (0, 1]$ and $\varepsilon < \frac{\rho}{C}$. We provide three methods to find a sufficient $\delta$ to define first order lattice as in lemma \ref{lem:main}.

For the first one, we assume that we have know a function that majorates $\Lambda(\sum_{n \geq 1}f_n)$.

\begin{prop} \label{prop : maj Lambda} \label{prop:locanalytic}
If $m_1$ is a real convex function above $\Lambda(\sum_{n \geq 1}f_n)$.
We define $m_2$ by extending $m_1$ with slope $2$.
Let $\delta$ be the abscissa of the smallest intersection of $m_2$ and the line defined by $y=-\log \varepsilon+ x,$ or $v_\infty$ if there is no such intersection.  Then $e^\delta$ is enough to define first order lattices as in lemma  \ref{lem:main}.
\end{prop}
\begin{proof}
$\delta$ is well-defined thanks to corollary \ref{cor : trunc series}, and $m_2 \geq \Lambda(\sum_{n \geq 2}f_n).$

It is sufficient to prove that for all $x \in B_{\overline{E}}(0,e^\delta),$ $\vert \sum_{n\geq 2} f_n(x) \vert \leq \varepsilon \vert x \vert.$

By comparing slopes, if $x \in B_{\overline{E}}(0,e^\delta),$ $u = \log \vert x \vert,$  $\Lambda( \sum_{n \geq 2}f_n)(u) \geq -\log \varepsilon+ u.$
With proposition \ref{prop : connexion sup Newton}, for any $m \geq 2,$ $-\log \vert f_m \vert +m u \geq -\log \varepsilon+ u.$

Therefore, for any $m \geq 2,$ $\vert f_m \vert  \vert x \vert^m \leq \varepsilon \vert x \vert.$
By ultrametricity,   $\vert \sum_{n\geq 2} f_n(x) \vert \leq \varepsilon \vert x \vert,$ which concludes the proof.
\end{proof}

For the second one, we only assume that we have majorations on the $\vert f_n \vert.$

\begin{prop}
Let $(M_n)_{n \geq 2}$ be real numbers such that for all $n \geq 2$, $\vert f_n \vert \leq M_n.$

If we denote by $\NP((M_{n+1})_{n \geq 1})$ the lower convex hull of the $(n, -\log M_{n+1})$ for $n \geq 1$, then $\delta$ such that $-\log \varepsilon=\NP((M_n)_{n \geq 2})^*(r)$ is enough to define first order lattices as in lemma  \ref{lem:main}.
\end{prop}
\begin{proof}
Same idea as proposition \ref{prop : maj Lambda}. One can define $g_n$ with $g_n (x) =\frac{f_n(x)}{\vert x \vert}$ and $g_n(0)=0$ and work with $\Lambda ( \sum g_n).$
\end{proof}

We finally provide a simple method when working with finite-dimensional vector spaces over $p$-adic fields.

\begin{prop} \label{prop : locanalyticfinitedim}
If $f$ is a locally analytic map $K^n \rightarrow K^m$, with $K$ a $p$-adic field, if $C,$ $\rho$ and $\varepsilon$ are as above. It is enough to work with the Newton polygon $NP_F$ of $F(t) = \sum_{k\geq 0}^{+\infty} \interleave f^{(k+1)}(0) \interleave  t^k \in \mathbb{R}[[t]]$ : let $\delta$ be such that $NP_F^* \left( \log_p \delta - \frac{1}{p-1} \right) \leq \log_p \varepsilon.$ Then $\delta$ is enough for lemma \ref{lem:main}. 
\end{prop}
\begin{proof}
Same as before, with $L_n=\frac{f^{(n)}(0)}{n!}$ and Legendre formula. The result follows \textit{mutatis mutandis}.
\end{proof}

\todo{Explain how Newton polygons considered above behave under
usual operations (sum, multiplication and composition of functions).}

%Using Lemma \ref{lem:reductionball}, one can extend Proposition 
%\ref{prop:locanalytic} to arbitrary lattices $H$ as follows. Given $H$
%a lattice in some $K$-Banach space $E$, we first define the two
%following quantities:
%\begin{itemize}
%\item $\max(H)$ is the smallest $r$ such that $H \subset B_E(r)$, and
%\item $\min(H)$ is the greatest $r$ such that $B_E(r) \subset H$.
%\end{itemize}
%
%\begin{cor}
%Let $U$ are $V$ be open subsets in $K$-Banach spaces $E$ and $F$ 
%respectively and $f : U \to V$ be a locally analytic function. Let 
%also $v_0$ be a point in $U$ such that $f'(v_0)$ is surjective. 
%Then a lattice $H$ in $E$ is a first order lattice for $f$ at $v_0$ as 
%soon as $\max(H)^2 < C_{f, v_0} \cdot \min(H)$ where $C_{f, v_0}$ is a 
%constant depending only on $f$ and $v_0$ which can be determined 
%explicitely from the Newton polygon of $f$ at $v_0$.
%\end{cor}
%
%\begin{proof}
%We apply Proposition \ref{prop:locanalytic} to the composite $f \circ 
%\iota_H$ where $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
%\Vert)$ is induced by identity map.
%\end{proof}

@@

\subsubsection*{Conclusion}

The theory developped in this Appendix is nice because it provides an 
abstract framework to deal with ultrametric precision in a quite general 
setting.
Nevertheless, as we have seen in the above examples, in practice we
generally carry out all calculations in charts and do not really care
about the ``real'' nature of the precision data, \emph{i.e.} lattices
in tangent spaces.

@@

\bibliographystyle{plain}
\bibliography{roebib/Biblio}

>>>>

We first note that a LU decomposition of a given matrix $M$ is unique
if we require further that the $L$-part of this decomposition is
unipotent (\emph{i.e.} all diagonal entries equal $1$). Let us write
$L(M) U(M)$ the LU factorization of $M$ satisfying the above extra
condition (when it exists). There exists Cramer-type formulae that
gives each individual entry of $L(M)$ and $U(M)$ in a closed form
(see \cite{caruso:12a}).
If $M$ is random square matrix over $\Zp$ of side $d$ whose every
coefficient is known up to precision $O(p^N)$, the two following
results are proved in \cite{caruso:12a}:
\begin{itemize}
\item using usual Gauss elimination and tracking $p$-adic precision 
step-by-step, the smallest precision on an entry of $L(M)$ is about
$O(p^{N - \frac{2d}{p-1}})$ on average;
\item computing $L(M)$ by evaluating Cramer-type formulae yields a
result whose every entry is known up to precision $O(p^{N - 2 \log_p 
d})$.
\end{itemize}
If $d$ is large compared to $p$, the second precision is much more
accurate than the first one. This shows once again that ``artifical''
losses of precision appear when one performs Gauss elimination by
tracking precision step-by-step.

@@

\begin{deftn} \label{deftn : diff}
Let $E$ and $F$ be two normed $K$-vector spaces, let $U$ be an open 
subset of $E$ and let $f : U \rightarrow F$ be a map. Then $f$ is called 
\emph{differentiable} at $v_0 \in U$ if there exists a continuous linear 
map $f'(v_0) : U \rightarrow W$ such that for any $\varepsilon >0$, 
there exists an open neighborhood $U_\varepsilon \subset U$ containing 
$v_0$ with
\[ 
\Vert f(v)-f(w)-f'(v_0) \cdot \left( v-w \right) \Vert 
\leq \varepsilon \Vert v-w \Vert. 
\]
for all $v, w \in U_\varepsilon$.  The linear map $f'(v_0)$ is called the \emph{differential} of $f$ at $v_0$.
\end{deftn}

\begin{rem}
This notion of differentiability is stronger than the usual notion with 
which we are familiar. It implies for instance the following: if $f$ is 
differentiable on $U$, then the function $x \mapsto f'(x)$ defined on 
$U$ is continuous. 
It is sometimes called \emph{strict differentiability} but, for this
article, we prefer a simplier terminology since we shall only use this
notion of differentiability.
\end{rem}

@@

\begin{proof} % Call it the Banach-Schauder theorem?
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, the open mapping theorem provides a $C>0$ such that 
$$B_F(1) \subset f'(0)(B_E(C)).$$
Let $\varepsilon>0$ be such that 
$\varepsilon C < \rho$, and choose $U_\varepsilon \subset E$ as in Definition 
\ref{deftn : diff}.  We may assume $U_\varepsilon = B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$. We suppose that $H$ is a lattice with 
$B^-_E(\rho r) \subset H \subset B_E(r).$
We seek to show that $f$ maps $H$ surjectively onto $f'(0) (H)$. We first prove 
that $f(H) \subset f'(0) (H)$. Suppose $x \in H$; we first show that $f(x) \in f'(0)(H)$.  By differentiability at 
$0$, $\Vert f(x)-f'(0)(x) \Vert \leq \varepsilon \Vert x \Vert $. 
Setting $y=f(x)-f'(0)(x)$, we have $\Vert y \Vert \leq 
\varepsilon r$. The definition of $C$ implies that $B_F(\varepsilon r) 
\subset \varepsilon f'(0) (B_E(\varepsilon rC))$. Thus there exists $x' 
\in \cdot B_E(\varepsilon r C)$ such that $f'(0) (x') =y$. Since 
$\varepsilon C < \rho$, we get $x' \in B^-_E(\rho r) \subset H$ and 
then $f(x)= f'(0) (x-x') \in f'(0) (H)$.

@@

We consider in this subsection the case where $f$ is locally analytic: 
under this extra assumption, we make explicit the constant $\delta$ 
which appears in Lemma \ref{lem:main}.
Locally analytic functions over finite dimensional $K$-vector spaces are 
defined in Schneider's book \cite {schneider:11a}, \S 6. His definition 
extends readily to $K$-Banach spaces as follows:

\begin{deftn}
\label{def:locanal}
Let $E$ and $F$ be $K$-Banach spaces. Let $U$ be an open subset of $E$
and let $x \in U$.
A function $f \col U \rightarrow F$ is said \emph{locally analytic} at 
$x$ if there exist an open subset $U_x \subset E$ and, for all $n \ge 
1$, a continuous $n$-linear map $L_n \col E^n \rightarrow F$ such that
$$f(x+h) = f(x) + \sum_{n \geq 1} L_n(h,\ldots,h)$$
for all $h$ with $x + h \in U_x$.
\end{deftn}

\begin{rmk}
A function $f$ which is locally analytic at $x$ is \emph{a fortiori} 
differentiable at $x$. Furthermore, in that case, the linear map $L_1$ 
appearing in the above definition is nothing but the derivative $f'(x)$.
\end{rmk}

From now, we assume that $K$ is algebraically closed. As in Definition 
\ref{def:locanal}, we consider two $K$-Banach spaces $E$ and $F$ and a 
family of continuous $n$-linear maps $L_n : E^n \to F$. For $n \ge 1$ 
and $h \in E$, we set $f_n(h) = L_n(h, \ldots, h)$ and
$$\Vert f_n \Vert = \sup_{h \in B_E(1)} \Vert f_n(h) \Vert.$$
When the series $\sum_n f_n(h)$ converges, we denote by $f(h)$ its sum;
we shall write $f = \sum_{n \geq 0} f_n$. We assume that $f$ is defined
in a neighborhood of $0$. Under this assumption, the datum of $f$ 
uniquely determines the $f_n$'s (it is a consequence of Propostion 
\ref{prop:lambdaNP} below).
To such a series $f$, we attach the function $\Lambda(f) : \R \to \R 
\cup \{+\infty\}$ defined by:
$$\begin{array}{rcll}
\Lambda(f)(v) & = & 
\log \big( \sup_{h \in B^-_E(e^v)} \Vert f(h) \Vert \big)
& \text{if } f \text{ is defined on } B^-_E(e^v) \\
& = & +\infty & \text{otherwise}
\end{array}$$
The following lemma is easy and left to the reader.

\begin{lem}
\label{lem:opLambda}
Let $f = \sum_{n \geq 0} f_n$ and $g = \sum_{n \geq 0} g_n$ be two 
series as above. Then:
\begin{align*}
\Lambda(f+g) &\leq \max (\Lambda(f),\Lambda(g)) \\
\Lambda(f \times g) &\leq \Lambda(f)+\Lambda(g) \\
\Lambda(f \circ g) &\leq \Lambda(f) \circ \Lambda(g)
\end{align*}
\end{lem}

\begin{rem}
\label{rem:opLambda}
Using Lemma \ref{lem:opLambda}, one can easily derive an upper 
bound of $\Lambda(f)$ from a formula describing $f$.
\end{rem}

The function $\Lambda(f)$ we have just defined is actually closely 
related to the Newton polygon of $f$. Recall that, by definition, the 
Newton polygon of $f$ is the convex hull in $\R^2$ of the points of 
coordinates $(n, - \log \lVert f_n \rVert )$ for $n \geq 0$ together 
with the extra point $(0, +\infty)$. We denote by $\NP(f) : \R \to \R 
\cup \{+\infty\}$ the convex function whose epigraph is the Newton 
polygon of $f$.

We recall that the Legendre transform of a \emph{convex} function 
$\varphi : \R \to \R \cup \{+\infty\}$ is the function $\varphi^\star : 
\R \to \R \cup \{+\infty\}$ defined by:
$$\textstyle \varphi^\star(v) = \sup_{u \in \R} \:
\big(uv - \varphi(u)\big), \qquad
\forall v \in \R.$$
One can check that the map $\varphi \mapsto \varphi^\star$ is an
involution (\emph{i.e.} $(\varphi^\star)^\star = \varphi$) and
reverses the order (\emph{i.e.} if $\varphi \leq \psi$, then
$\varphi^\star \geq \psi^\star$). 
%Furthermore, if $\varphi$ is
%piecewise affine (and convex), the slopes of $\varphi^\star$ are
%exactly the points where $\varphi$ is not differentiable and \emph{vice
%versa}.

%\begin{prop} \label{prop : conjugate Newton}
%Considered as a function on $\R$, $\NP(f)^*$ is defined over $\left] -\infty,v_{\infty} \right[$, though it may be defined for larger values.
%\end{prop}
%
%We can interpret the function $\Lambda(f)$ in terms of the conjugate Newton polygon:

\begin{prop} \label{prop:lambdaNP}
Keeping the above notation, we have $\Lambda (f) = \NP(f)^\star$.
\end{prop}

\begin{proof}
Let $v \in \R$. We assume first that $\NP(f)^\star(v) = +\infty$. Then, 
using a multilinear version of Banach-Steinhaus Theorem, we can prove 
that $\sum_{n \geq 0} f_n$ does not converge on $B_E^-(e^v)$ and 
therefore that $\Lambda(f)(v) = +\infty$ as well.

We now suppose that $\NP(f)^\star(v)$ is finite. To begin with, we 
assume in addition that $v$ is not a slope of $\NP(f)$. We set $u = 
\NP(f)^*(v)$. The function $m \mapsto \NP(f)(m) - vm + u$ has the 
following properties: (1)~it is piecewise affine and everywhere 
nonnegative, (2)~it does not admit $0$ as a slope and (3)~it vanishes at 
$x = n$ for some integer $n$ and $u = vn + \log \Vert f_n \Vert$.
We deduce from this that there exists $c>0$ such that $vm-u \leq - \log 
\lVert f_m \rVert - c \cdot \vert n-m \vert$ for any $m\geq 0$.
Since $vm-u=vm-vn-\log \lVert f_n \rVert$, we get $-vn-\log \lVert 
f_n \rVert + c \cdot \vert n-m \vert \leq -vm-\log \lVert f_m \rVert$.
Therefore, for any $x \in  B_E(e^v)$ and $m \geq 0,$ we have
$$\Vert f_m (x) \Vert \leq e^{-c \cdot \vert n-m \vert } \cdot
\lVert f_n \rVert \cdot e^{vn}< \lVert f_n \rVert \cdot e^{vn}.$$
Thus, the series $\sum_{m \geq 0} f_m(x)$ converges and $\Vert f(x) 
\Vert \leq \Vert f_n \Vert \cdot e^{vn}$. We then get:
\begin{equation}
\label{eq:lowerLambda}
\Lambda (f)(v) \leq \log \big( \Vert f_n \Vert e^{vn} \big)
= vn + \log \Vert f_n \Vert = u.
\end{equation}
On the other hand, it follows from the definition of $\lVert f_n \rVert$ 
that there exists a sequence $(x_i)_{i\geq 0}$ in $B^-_E(e^v)$ such that 
$\lim \Vert f_n (x_i) \Vert= \lVert f_n \rVert \cdot e^{vn}$. Since 
furthermore $\Vert f_m (x_i) \Vert \leq e^{-c \cdot \vert n-m \vert} 
\Vert f_n \Vert \cdot e^{vn}$ for all $i$, we get $\Vert f_m (x_i) \Vert 
< \Vert f_n (x_i) \Vert$ for $i$ large enough. For these $i$, we then 
have $\Vert f(x_i) \Vert = \Vert f_n (x_i) \Vert$. Passing to the limit 
on $i$, we find $\Lambda (f)(v) \geq u$. Comparing with 
Eq.~\eqref{eq:lowerLambda}, we get $\Lambda (f)(v) = u = 
\NP(f)^\star(v)$.

We have then proved that the functions $\Lambda(f)$ and $\NP(f)^\star$ 
coincide expect maybe on the set consisting of slopes of $\NP(f)$. But 
we remark that this set is countable and that both functions $\Lambda 
(f)$ and $\NP(f)^\star$ are left continuous. This is enough to ensure 
that $\Lambda (f)$ and $\NP(f)^\star$ coincide everywhere.
\end{proof}

\begin{rem}
It follows from Proposition \ref{prop:lambdaNP} that $\Lambda(f)$ is a
convex function (which was not clear at all \emph{a priori}).
\end{rem}

We now study the effect of truncation on series: given $f$ as above
and a nonnegative integer $n_0$, we set:
$$f_{\geq n_0} = \sum_{n \geq n_0} f_n = f - (f_0 + f_1 + \cdots +
f_{n_0 - 1}).$$
On the other hand, given a convex function $\varphi : \R \to \R \cup
\{+\infty\}$ and a real number $v$, we define $\varphi_{\geq v} : \R \to 
\R \cup \{\pm \infty\}$ as the highest convex function such that 
$\varphi_{\geq v} \leq \varphi$ and the function $x \mapsto 
\varphi_{\geq v}(x) - v x$ is not decreasing. Concretely, we have:
$$\varphi_{\geq v}(x) = \inf_{y \geq 0} \, \big(\varphi(x + y) - v y 
\big).$$
When $v$ is fixed, the construction $\varphi \mapsto \varphi_{\geq v}$
is nondecreasing: if $\varphi$ and $\psi$ are two convex functions such
that $\varphi \leq \psi$, we deduce $\varphi_{\geq v} \leq \psi_{\geq v}$.

\begin{prop} \label{prop:trunc}
With the above notations, we have $\Lambda(f_{\geq n_0}) \leq
\Lambda(f)_{\geq n_0}$ for all $n_0 \in \mathbb N$.
\end{prop}

\begin{proof}
It follows easily from Proposition \ref{prop:lambdaNP} and the fact that 
the slopes of the Legendre transform of a convex piecewise affine 
function $f$ are exactly the points where $f$ is not differentiable.
\end{proof}

%Now, let $E$ and $F$ be two complete normed $K$-vector spaces, and let $f$ and the $L_k$'s be as in definition \ref{deftn:analyticNP}. We assume $L_1$ is surjective. We define, the $\overline{L_k},$ $\overline{E}$ and $\overline{F}$ by extending scalars to $\overline{K}.$ We define for any $n \geq 1$, $f_n(x)=\overline{L_n}(x,\dots,x)$ for any $x \in \overline{E}.$ 
%Let $C$ be given by the open mapping theorem for $L_1$, $\rho \in (0, 1]$ and $\varepsilon < \frac{\rho}{C}$. We provide three methods to find a sufficient $\delta$ to define first order lattice as in lemma \ref{lem:main}.
%
%For the first one, we assume that we have know a function that majorates $\Lambda(\sum_{n \geq 1}f_n)$.

\begin{prop}
\label{prop:locanalytic}
Let $f = \sum_{n \geq 0} f_n$ be a function as above.
Let $C$ be a positive real number satisfying 
$B_F(1) \subset f_1(B_E(C))$.
Let $\rho \in (0, 1]$ and $\nu$ be a real number such that 
\begin{equation}
\label{eq:locanalytic}
\Lambda(f)_{\geq 2} (\nu) < \nu + \log \Big( \frac \rho C \Big).
\end{equation}
Then the conclusion of Lemma \ref{lem:main} holds with $\delta = e^\nu$.
\end{prop}

\begin{rem}
On a neighborhood of $-\infty$, the function $x \mapsto \Lambda(f)_{\geq 
2} (x) - x$ is affine of slope $1$. This implies that, for all $\rho \in
(0,1]$, there exists $\nu$ satisfying \eqref{eq:locanalytic}. Moreover,
if $\rho$ is close enough to $0$, then one can take $\delta = e^\nu$ as a 
linear function of $\rho$.
\end{rem}

\begin{proof}
Pick $\varepsilon$ in the interval $(e^{\Lambda(f)_{\geq 2} (\nu) - \nu}, 
\frac \rho C)$. Going back to the proof of Lemma \ref{lem:main}, we 
observe that it is enough to prove that
\begin{equation}
\label{eq:boundqueue} 
\Vert f_{\geq 2}(x) \Vert \leq \varepsilon \cdot \Vert x \Vert.
\end{equation}
for all $x \in B_E(\delta)$.
This follows directly from Proposition \ref{prop:lambdaNP} and Proposition
\ref{prop:trunc}.
\end{proof}

\begin{rem}
It follows from the proof that Proposition \ref{prop:locanalytic} is 
still valid if $K$ is not assumed to be algebraically closed. Indeed, 
the functions $f_n$ --- and then $f$ also --- extend to an algebraic
closure $\bar K$ of $K$ and Eq.~\eqref{eq:boundqueue} holds over $\bar 
K$, which is enough to conclude.
\end{rem}

\begin{cor}
We keep the notations of Proposition \ref{prop:locanalytic} and consider 
in addition a sequence $(M_n)_{n \geq 2}$ such that $\Vert f_n \Vert 
\leq M_n$ for all $n \geq 2$. Let $\NP(M_n)$ denote the convex function
whose epigraph is the convex hull in $\R^2$ of the points of coordinates
$(n, -\log M_n)$ for $n \geq 2$ together with the extra point $(0,
+\infty)$.

Let $\rho \in (0, 1]$ and $\nu$ be a real number such that 
$$NP(M_n)^\star (\nu) < \nu + \log \Big( \frac \rho C \Big).$$
Then the conclusion of Lemma \ref{lem:main} holds with $\delta = e^\nu$.
\end{cor}

\begin{rem}
If $K$ has characteristic $0$ and the vector spaces $E$ and $F$ are
finite dimensional, then the $M_n$'s defined by
$$M_n = \frac 1 {|n!|} \cdot 
\sup_{\substack{1 \leq i \leq \dim E \\ \vert \underline n
\vert = n}} \,
\Big\Vert \frac{\partial^n f_i}{\partial x^{\underline n}}(0) 
\Big\Vert$$
do the job. Here $f_i$ denotes the $i$-th coordinate of $f$, the
notation $\underline n$ refers to a tuple of $(\dim F)$ nonnegative
integers and $\vert \underline n \vert$ is the sum of the coordinates
of $\underline n$.
\end{rem}

@@

%\subsubsection*{Conclusion}
%
%The theory developped in this Appendix is nice because it provides an 
%abstract framework to deal with ultrametric precision in a quite general 
%setting.
%Nevertheless, as we have seen in the above examples, in practice we
%generally carry out all calculations in charts and do not really care
%about the ``real'' nature of the precision data, \emph{i.e.} lattices
%in tangent spaces.

@@

\bibliographystyle{plain}
\bibliography{Biblio}
